---
title: "R Notebook"
output: html_notebook
---

I will be writing down codes I found useful in this notebook :)

Add a new chunk by pressing *Ctrl+Alt+I*.
Run whole code chunk *Ctrl+Shift+Enter*.
select to line start *Alt+Shift+Left*.
Go to next chunk *Ctrl+PgDown*.

bullet lists
- hey  # here, you have to add two spaces after "hey"


Packages
```{r}
library(devtools)
library(dplyr)
library(ggplot2)
library(statsr)
library(shiny)
library(tidyverse)
library(vcd)
library(ggpubr)
library(knitr)
library(kableExtra)
library(GGally)
```


```{r}
getwd() # checking where your workind directory is. if need to change wd, then use setwd("URL of folder")
ls() # gives all the list of objects in the environment
```

##Basics
####Vectors
#####sequence
```{r}
20**2 #squared
3**3
4**3
sqrt(9) # this gives you 3
log(10)


d <- c(1:5)
d[d %in% 1:3] <- 0
d
#names
z <- 1:3
names(z) <- c("Y", "U", "I")
names(z)

#round
bo <- 0.3567
round(bo, digits = 2)

# vector can contain only same class of objects

fac <- factor(c("1 star", "2 stars", "3 stars", "3 stars", "4 stars", "5 stars"))
levels(fac) # this gives how many values there are in the vector
nlevels(fac) # this gives how many levels there are
length(fac)
fac[5] # this gives the 5th value in the vector (which is 4 stars in this case) 
fac[2:5] # this gives the 2nd to 5th values in the vector
fac[-1] # this gives values of the vector except for the 1st value
# sequence
1:14 # this can be done also with the `seq` as below
seq(1, 14) # this gives all the number from 1 to 14
seq(1, 14, by = 2) # this gives odd numbers within the range of 1 to 14
seq(2, 14, by = 2) # this gives even numbers within the range of 2 to 14
```

####Matrices
```{r}
matrix1 <- matrix(data = fac, nrow = 3, ncol = 3)
matrix1
matrixbyrow <- matrix(data = fac, nrow = 3, ncol = 3, byrow = TRUE)
matrixbyrow # this orders values in by row
int <- 1:16
matrix2 <- matrix(int, nrow = 4, ncol = 4)
matrix2
matrix2[3,3] #the value in 3rd row and 3rd column of the matrix2
matrix2[3,] # all the values in 3rd row
matrix2[,3] # all the values in 3rd column
matrix2[1:3,1:3] # all the values first 3rd rows and 3rd columns
matrix2[-4, -4] # this gives the same outcome as the one above
matrix2[c(1,3),] # values in 1st and 3rd rows
matrix2[,c(1,3)] # values in 1st and 3rd columns

#dimnames
dimnames(m) <- list(c("a","b","c","d","e"), c("A","B"))
m

k <- seq(from = 100, to = 1000, by = 20)
str(k)

#complete.cases() this seems quite useful :)
b <- c(1,2,NA,4,5,6)
c <- c("a",NA,NA,"d",NA,"f")
good <- complete.cases(b,c)
good
b[good]
c[good]

```

##### vectorized operations
```{r}
# vectorized operation see how the data parallels
e <- 1:4; f <-6:9
e + f
e*f

g <- matrix(1:4,2,2); h <- matrix(rep(10,4),2,2)
g*h
g%*%h
```

```{r}
# counting the number of NAs

hw1_data <- read.csv(file = "C:/Users/ynaru/Desktop/R coursera/Statistics with R/hw1_data.csv",header = T)
table(is.na(hw1_data$Ozone))
mean(hw1_data$Ozone, na.rm = T)
hw1_data2 <- subset(hw1_data, Ozone >31 & Temp > 90)
mean(hw1_data2$Solar.R)
class(hw1_data$Month)
mean(subset(hw1_data, Month == 6)$Temp)
max(subset(hw1_data, Month == 5)$Ozone, na.rm = T)
```



####Data frame
#####rbind, cbind, [], c(), &, |
```{r}
names <- c("Bob", "Nancy", "Kevin", "Anne", "David", "Lucy")
score <- c(88, 79, 65, 94, 77, 82)
grade <- data.frame(names, score)

grade$score 
grade[,2] # these two are the same. calls second column of the dataframe

rich <- data.frame(name = c("Jeff Bezos", "Bernard Arnault", "Elon Musk", "Bill Gates", "Mark Zuckerberg"), 
                   worth_billion = c(188.3, 188.0, 151.3, 126.5, 118.5), 
                   age = c(57, 72, 49, 65, 37), 
                   source = c("Amazon", "LVMH", "Tesla, SpaceX", "Microsoft", "Facebook"),
                   country = factor(c("United States", "France", "United States", "United States", "United States")))

twomore <- data.frame(name = c("Warren Buffett", "Larry Page"), 
                      worth_billion = c(109.4, 103.0), 
                      age = c(90, 48), 
                      source = c("Berkshire Hathaway", "Google"), 
                      country = factor(c("United States", "United States")))
rich <- rbind(rich, twomore) # this adds new rows

education <- factor(c("Princeton University", "Ecole Polytechnique", "University of Pennsylvania", "Harvard University", "Harvard University", "Columbia Business School", "Stanford University")) # create new vectors
rich <- cbind(rich, education) # this combines new column
str(rich)
head(rich, 3) #calls first three rows of the dataframe
rich[1:5, "source"] # select 1 to 5th rows of the column name "source"
rich[3:7, c("age","education")] # select 3 to 7th rows of the column "age" and "education"
rich[rich$country == "United States", c("name", "worth_billion", "education")] # select people from United States and show values of name, income and education
rich[rich$country == "United States" & rich$education == "Harvard University",] # select people from United States and also graduated from Harvard University
rich[rich$country == "France" | rich$age < 60, c("name", "country", "age")] # select people from France OR age under 60 and show name, country and age
```

#####correcting, ordering, finding names
United States to US
age order
```{r}

#correction
rich2 <- rich
str(rich2)
rich2$country <- as.character(rich2$country) # first i had to change country to character and then assign US. 
class(rich2$country)
rich2[rich2$country == "United States", c("country")] <- "US"
rich2 # worked!! :)

#ordering and arranging
ordered(rich2$age) # show the order of age
rich3 <-arrange(rich2, age) # to order the data based on age
rich3
rich4 <- arrange(rich2, desc(age))
rich4 # arrange dataframe with descending order based on age

#searching
grep("edu", names(rich), value = TRUE) # searches if there are any variable name with something similar to"edu" in dataframe rich.

```

####Creating new column based on other column
```{r}
#case_when
mtcars2 <- mtcars
mtcars2 <- mtcars2 %>%
  mutate(displevels = 
           case_when(disp <= 200 ~ "Low",
                     disp <= 300 ~ "Medium",
                     disp <= 400 ~ "High",
                     disp > 401 ~ "Really high"))
```



####filtering
```{r}
#filtering

grade
gender <- c("M","F","M","F","M","F")
major <- c("Science", "Literature", "History", "Science", "Science","Architecture")
GPA <- c(2.8,3.2,3.7,2.7,3.0,3.5)
grade <- cbind(grade, GPA)
grade
grade %>%
  filter(gender == "M", major == "Science", GPA >= 3)

grade %>%
  filter(GPA %in% c(3, 3.7)) %>%
  group_by(GPA) # only gives observations where GPA is either 3 or 3.7
grade %>%
  filter(major %in% c("Science", "History")) %>%
  group_by(GPA) %>%
  summarise(count = n()) %>%
  filter(GPA < 3) # first summarize by GPA with filtering the observations only with people from science and history, then filter the results by setting GPA range as smaller than 3.
```





####List
list is something you can store many things (matrices, vectors in a same place)
list is special in a sense that it can store different kinds of class of objects (not like vector, created with c())
```{r}
list <- list(matrix2, grade, rich)
list
list[[2]] # calls second object "grade" from the list
list[[2]][1,] # calls 1st row of the second object
list2 <- list(2, "s", F)
list2[[1]]

```







##Plots


```{r}

```


##Statistical analysis

#####Basics
mean, median, mode, range, quantile, IQR
```{r}
xdata <- c(1,3,4,5,6,6,6,8,9,12,NA)

as.numeric(names(which.max(table(xdata)))) # getting mode
median(xdata, na.rm = TRUE)
mean(xdata, na.rm = TRUE)
xdata2 <- c(xdata[-10], 35)
xdata2 # removing 10th value (12) and adding 35 at the last
mean(xdata2, na.rm = TRUE)
median(xdata2, na.rm = TRUE)
range(xdata2, na.rm = TRUE)

quantile(xdata2, probs = c(0.25, 0.75), na.rm = TRUE) # 25% and 75% value
quantile(xdata2, prob = seq(0,1,0.2), na.rm = TRUE) # all the percentile from 0% to 100% in 20% interval :)
IQR(xdata2, na.rm = TRUE)
summary(xdata2)
```

tapply
```{r}
tapply(rich2$worth_billion, rich2$country, mean) # mean of income grouped by country
tapply(rich2$worth_billion, rich2$age<50, mean) # mean of income grouped by age (i.e. if age less than 50 TRUE)
tapply(rich2$age, list(rich2$country, rich2$education), mean, na.rm = TRUE) # although in this example the outcome does not appear that interesting, it basically gives age mean grouped by two variables
```


####Simulations
```{r}
coin_outcomes <- c("heads", "tails")
sample(coin_outcomes, size = 1, replace = TRUE) # which one we get when we flip a coin. you can do it a few times and get different outcome :)
sim_fair_coin <- sample(coin_outcomes, size = 100, replace = TRUE)
head(sim_fair_coin, 20)
table(sim_fair_coin) # you can see that it got 48 heads and 52 tails
sim_unfair_coin <- sample(coin_outcomes, size = 100, replace = TRUE, prob = c(0.1, 0.9))
table(sim_unfair_coin) # you can see that in this case, we got only 15 heads because I specified the probability of success(heads) as 0.1 probability.


```






####Probability
#####Normal distribution
pnorm, qnorm, rnorm, dnorm, replicate
```{r}
#pnorm = probability of observing a particular value 
pnorm(7, 11.2, 2.8) # pnorm(observation, mean, sd). observing 7 when the mean is 11.28 and sd is 2.8
pnorm(15, 11.2, 2.8, lower.tail = FALSE) # this calculates upper value since 15 is more than the mean

#qnorm = what is the value of particular percentile, i.e. quantile?
qnorm(0.2, 11.2, 2.8) # qnorm(proportion, mean, sd). what is the value at 20% when the mean is 11.28 and sd is 2.8.
qnorm(0.2, 11.2, 2.8, lower.tail = FALSE) # upper = 80%
qnorm(0.8, 11.2, 2.8) # this is the same as the above line

#rnorm = creates random data
males <- rnorm(15, 10.3, 2.6) # (number of values, mean, sd)
females <- rnorm(15, 6.4, 3.8)
both <- c(males, females)
both
sex <- c(rep("M", 15), rep("F", 15)) # replicates 15 M and F for each
sex
range_sex <- cbind(both, sex)
range_sex

# dnorm = probability of distribution function
sample.distribution <- 50:150
distribution <- dnorm(sample.distribution, 100, 15) # how the distribution will look like when there are 101 observations with mean 100 and sd 15
iq.df <- data.frame(sample.distribution, distribution)
ggplot(data = iq.df, aes(x=sample.distribution, y=distribution)) +
  geom_point()
sum(iq.df$distribution) # almost 100%
```



#####Bernoulli trial and Binomial distribution
success/failure condition, n choose k
```{r}
dbinom(5, 6, 0.5) # (k(number of success), n(number of trial), p(probability of success) )

0.5**6*6 # cooool this gives the same value. 


dbinom(3, 6, 1/6) # e.g., getting 6 three times in six trials

success <- (1/6)**3
failure <- (5/6)**3
scenario <- success*failure
ways <- choose(6,3) # this is n choose k (possible order of getting k successes in n trials)
scenario*ways # this gives the same result. Or simply = (1/6)**3*(5/6)**3*20 


sum(dbinom(0:4, size = 20, p = 0.5)) # this gives the p-value of, if in fact the probability of getting pregnant is equally the same (0.5), how probable to get k = 4 or less (number of women who get pregnant), when the size is 20.
dbinom(0:4, size = 20, p = 0.5) # if did not use sum function, it gives all the p-value for pregnant 0 to 4 women.


```

Ex: What would be the probability of getting 1 or more yellow M&M in a bag of 5 M&M’s? (there are 5 possible color including yellow)
```{r}
prob = dbinom(1, 5, 0.1) + dbinom(2, 5, 0.1) + dbinom(3, 5, 0.1) + dbinom(4, 5, 0.1) + dbinom(5, 5, 0.1)

# or, use at least way of thinking

prob = 1 - dbinom(0, 5, 0.1)

prob
```





####Bayesian inference
```{r}
# let's see the probability of 4 women getting pregnant among the 20 women who have been assigned morning after pill.

p <- seq(from = 0.1, to = 0.9, by = 0.1)
prior <- c(rep(0.06, 4), 0.52, rep(0.06, 4))
likelihood <- dbinom(4, 20, prob = p)
likelihood_round <- round(likelihood, digits = 4) # rounding the decimal
prior_tab <- rbind(prior_tab, likelihood_round)
prior_tab

numerator <- prior*likelihood_round
denominator <- sum(numerator)
posterior <- numerator/denominator
posterior


```


### Conjugate family
```{r}
# gamma(4) = (4-1)! = 3*2*1 = 6
gamma(4)


# Poisson ‘Pois(k = 1, lambda = 2)‘
dpois(1, lambda = 2)

# Beta ‘Beta(x = 0.5, alpha = 2, beta = 3)‘
dbeta(0.5, shape1 = 2, shape2 = 3)

# Gamma ‘Gamma(x = 1, k = 2, theta = 3)‘
dgamma(1, shape = 2, scale = 3)

# Gamma ‘Gamma(x = 1, alpha = 2, beta = 3)
dgamma(1, shape = 2, rate = 3)

```


```{r}
dplyr::select(samples, mu, y_pred) %>% # get credible interval for both mu and y_pred. this line may not run though
  map(quantile, probs=c(0.025, 0.50, 0.975)) # map to get quantiles for both 
```


